Distributed Topology Discovery in Self-Assembled Nano Network-On-Chip

### 2014-09-11 PAPER/introduction.tex
Changes suggested by Andrew

Exploring long term alternatives to the CMOS technology is gaining
more and more relevance as the scaling trend of such devices keeps
introducing new challenges: power density, defect tolerance, testing
costs and wire delays are only a few of the many critical aspects
involved~\cite{itrs13}. 
While software parallelism and multicore
approaches~\cite{horowitz2004,powell2009} are partially mitigating the impact of such
constraints on performances, it is likely that a growing computing demand will
eventually need even more radical architectural modifications and 
paradigms to address the Computer Design challenges of the next
decades.

In the last years, self-assembled nanoscale
architectures~\cite{yan2003} emerged as a promising technology due
their tera/peta scale of integration, defect tolerance and huge
potential computing capabilities. These technologies are certainly
still at their early stage of development, however different
laboratory demos and proof-of-concepts have been
presented~\cite{patwardhan2004, patwardhan2006, patwardhan2006_1,
pistol2009}. For a complete survey on self-assembled architectures and
how they compare against classical many-core systems in terms of
 technologies, fault tolerance, performance and software tools see
 also~\cite{Future}.
%, patwardhan2006
The main idea behind this approach is exploiting the physical
regularity and stability of DNA structures in order to create a
scaffold onto which nano-devices (e.g. nanowires and
CNFETs~\cite{bachtold2001, tans1998, cui2001}) can be attached. This
can be achieved by designing appropriate complementary DNA tags for
each terminal to be placed, so that a nano device will be attached
only where its own DNA tag matches a complementary tag on the DNA grid
scaffold (see Figure~\ref{fig:nana}-a).  Of course, a detailed
description of the chemical properties involved is far beyond the
scope of this paper (see also~\cite{braun1998, seeman1999}).

In this work, we present \disr{} (Distributed Segment Routing), a distributed
approach focused on addressing the three main challenges that this new fabrication process introduces
in nanoscale Network-on-Chip Design: \emph{(i) limited node complexity}, \emph{(ii) large scale
randomness} and \emph{(iii) high defect rates}.  
The \emph{limited node complexity} aspect is directly
related to the use of complementary DNA tags in order to place circuit
components. The traditional CMOS process introduces complexity 
with larger photolithography masks: more complex (larger) circuits
will only require larger masks. Conversely, a self-assembly process achieves
complexity by increasing the number of unique DNA tags, since having more
different tags means having more control on component placement. Ideally, by specifying a single
and unique tag for each nano device terminal, we could exactly
choose where each component would be placed in the design, but the number of DNA
symbols forming the DNA sequence is limited (sequence of 4 nucleobases
G,A,T,C) and so creating many different tags (of a prefixed length) would
mean make them more similar to each other, increasing the probability
of incorrect/partial matching. To avoid this problem 
the number of unique tags must be limited, this limits also the
complexity available at each node.
In this work a budget of about $10,000$ CNFETs per node
has been assumed, as estimated in~\cite{liu_jetcs}.
\emph{Large scale randomness} is the other fundamental condition of
self-assembled technology: DNA tags allow control placement inside the
node grid, but there is no control over the placement of these grids
in the whole network. As a consequence, other typical
properties of regular networks cannot be guaranteed, e.g. being
connected to a fixed number of neighbors, having a determined
orientation and so on.
Finally, \emph{defect rates}: while they are very hardly tolerated in mask
based top-down design, the same nature of a bottom-up self-assembly
 process cannot assume such a deterministic device placement
process, thus defect tolerance is more a design requirement
than an exception to be avoided.

>>>>

Exploring long-term alternatives to the CMOS technology is gaining
more and more relevance as the scaling trend of such devices keeps
introducing new challenges: power density, defect tolerance, testing
costs and wire delays are only a few of the many critical aspects
involved~\cite{itrs13}. 
While software parallelism and multicore
approaches~\cite{Kavi2012243,Sibai2011958} are partially mitigating the impact of such
%approaches~\cite{horowitz2004,powell2009} are partially mitigating the impact of such
constraints on performances, it is likely that a growing computing demand will
eventually need even more radical architectural modifications and 
paradigm shifts to address the Computer Design challenges of the upcoming
decades.

In recent years, self-assembled nanoscale
architectures~\cite{yan2003} emerged as a promising technology due
their tera/peta scale of integration, defect tolerance and huge
potential computing capabilities. These technologies are certainly
still at their early stage of development; however, different
laboratory demos and proofs-of-concept have been
presented~\cite{patwardhan2006_1, pistol2009}. For a complete survey on self-assembled architectures and
%presented~\cite{patwardhan2004, patwardhan2006, patwardhan2006_1, pistol2009}. For a complete survey on self-assembled architectures and
how they compare against classical many-core systems in terms of
 technologies, fault tolerance, performance and software tools see
 also~\cite{Future}.
%, patwardhan2006
The main idea behind this approach is exploiting the physical
regularity and stability of DNA structures in order to create a
scaffold onto which nano-devices (e.g. nanowires and
CNFETs~\cite{bachtold2001, cui2001}) can be attached. This
%CNFETs~\cite{bachtold2001, tans1998, cui2001}) can be attached. This
can be achieved by designing appropriate complementary DNA tags for
each terminal to be placed, so that a nano device will be attached
only where its own DNA tag matches a complementary tag on the DNA grid
scaffold (see Figure~\ref{fig:nana}-a).  Of course, a detailed
description of the chemical properties involved is far beyond the
scope of this paper (see also~\cite{seeman1999}).

In this work, we present \disr{} (Distributed Segment Routing), a distributed
approach focused on addressing the three main challenges that this new fabrication process introduces
in nanoscale Network-on-Chip Design: \emph{(i) limited node complexity}, \emph{(ii) large scale
randomness} and \emph{(iii) high defect rates}.  
The \emph{limited node complexity} aspect is directly
related to the use of complementary DNA tags in order to place circuit
components. The traditional CMOS process introduces complexity 
with larger photolithography masks: more complex (larger) circuits
will only require larger masks. Conversely, a self-assembly process achieves
complexity by increasing the number of unique DNA tags, since having more
different tags means having more control on component placement. Ideally, by specifying a single
and unique tag for each nano device terminal, we could exactly
choose where each component would be placed in the design, but the number of DNA
symbols forming the DNA sequence is limited (sequence of 4 nucleobases
G,A,T,C) and so creating many different tags (of a predetermined length) would
mean making them more similar to each other, increasing the probability
of incorrect/partial matching. To avoid this problem 
the number of unique tags must be limited, which limits also the
complexity available at each node.
In this work, a budget of about $10,000$ CNFETs per node
has been assumed, as estimated in~\cite{liu_jetcs}.
\emph{Large scale randomness} is the other fundamental condition of
self-assembled technology: DNA tags allow controlled placement inside the
node grid, but there is no control over the placement of these grids
in the whole network. As a consequence, other typical
properties of regular networks cannot be guaranteed, e.g. being
connected to a fixed number of neighbors, having a determined
orientation and so on.
Finally, \emph{defect rates}: while they are hardly tolerated in 
mask-based top-down design, the same nature of a bottom-up self-assembly
 process cannot assume such a deterministic device placement
process, thus defect tolerance is more a design requirement
than an exception to be avoided.

### 2014-09-11 PAPER/paper.tex
Changes suggested by Andrew

\begin{abstract}
In this paper we present \disr{}, a distributed approach to topology
discovery and defect mapping in a self-assembled nano
network-on-chip. The main aim is achieving the already-proven
properties of segment-based deadlock freedom with neither requiring
a topology graph as input, nor needing a centralized algorithm to
configure network paths.  After introducing the conceptual elements
and the execution model of \disr{}, we show how the open-source Nanoxim
platform has been used to evaluate the proposed approach in the
process of discovering irregular network topology while establishing
network segments. Comparison against a tree-based approach shows how
\disr{} still preserves some important properties (coverage, defect
tolerance, scalability) while avoiding resource hungry solutions such
as virtual channels and hardware redundancy. Finally, we propose a
gate-level hardware implementation of the required control logic and
storage for \disr{}, demonstrating a relatively acceptable impact
ranging from 10 to about 20\% of the budget of transistors available
for each node.
\end{abstract}

@@

\section{Conclusions}
In this work we presented \disr{}, an initial attempt of achieving a segment based
topology discovery into distributed self-assembled nanoscale scenario. We
demonstrated how \disr{} can accomplish the aim of finding a segment
coverage of the network without requiring any centralized approach
that would request the graph topology as input. A draft hardware
implementation has been presented to evaluate the impact on the limited
node size typical of the assumed scenario. From an high-level
perspective future works will focus on
investigating \disr{} resulting network in order to support the
execution of massively parallel applications, while keeping into
developing a detailed low level hardware implementation on
DNA grid using the appropriate nano devices models.

>>>>

\begin{abstract}
In this paper, we present \disr{}, a distributed approach to topology
discovery and defect mapping in a self-assembled nano
network-on-chip. The main aim is to achieve the already-proven
properties of segment-based deadlock freedom requiring neither 
a topology graph as input, nor a centralized algorithm to
configure network paths.  After introducing the conceptual elements
and the execution model of \disr{}, we show how the open-source Nanoxim
platform has been used to evaluate the proposed approach in the
process of discovering irregular network topology while establishing
network segments. Comparison against a tree-based approach shows how
\disr{} still preserves some important properties (coverage, defect
tolerance, scalability) while avoiding resource hungry solutions such
as virtual channels and hardware redundancy. Finally, we propose a
gate-level hardware implementation of the required control logic and
storage for \disr{}, demonstrating a relatively acceptable impact
ranging from 10 to about 20\% of the budget of transistors available
for each node.
\end{abstract}

@@

\section{Conclusions}
In this work, we presented \disr{}, an initial attempt of achieving a segment based
topology discovery in a distributed self-assembled nanoscale scenario. We
demonstrated how \disr{} can accomplish the aim of finding a segment
coverage of the network without requiring any centralized approach
that would request the graph topology as input. A draft hardware
implementation has been presented to evaluate the impact on the limited
node size typical of the assumed scenario. From a high-level
perspective, future works will focus on
investigating \disr{} resulting networks in order to support the
execution of massively parallel applications, while continuing to
develop a detailed low level hardware implementation on a
DNA grid using the appropriate nano device models.
